---
title: 'Homework 4: Original Homework'
author: "Bhavya Deepti Vadavalli"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

```{r preliminaries}
library(curl)
library(ggplot2)
library(lmodel2)
```

## Question 1: 

Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:

1. Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().

2. When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().

3. The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

4. The function should contain a check for the rules of thumb we have talked about (n∗p>5
 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.
 
5. The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.

Note: To whoever is reviewing this, I'm so sorry. 

```{r question1}

z.prop.test <- function(p1,n1,p2,n2,p0,alternative,conf.level){
  alternative <- "two.sided"
  conf.level <- 0.95
  if(missing(p2|n2)){
        p2 <- NULL
        n2 <- NULL
        #going to one-sample z prop test
        z <- (p1 - p0)/sqrt(p0 * (1 - p0)/n1) 
        if (z > 0) {
            pval <- 2 * pnorm(z, lower.tail = FALSE)
          }
          if (z < 0) {
            pval <- 2 * pnorm(z, lower.tail = TRUE)
          }
    
      if (alternative == "greater") {
        p <- pnorm(z, lower.tail = FALSE)
      }
      if (alternative == "less") {
        p <- pnorm(z, lower.tail = TRUE)
      }
     
    #values for confidence interval   
    lower <- p1 - qnorm(conf.level+0.5*(1-conf.level)) * sqrt(p1 * (1 - p1)/n1)
    upper <- p1 + qnorm(conf.level+0.5*(1-conf.level)) * sqrt(p1 * (1 - p1)/n1)
    ci <- c(lower, upper)
    
    #returning the results: 
    return("one-sided z prop-test")
    CI <- conf.level
    CI
    Alternative <- alternative
    Alternative
    z-value <- as.numeric(z)
    p-value <- as.numeric(p) #I will make this more efficient if I have time
    
  } else{
   #two sample z-prop test with alternative = "two.sided"
    pstar <- ((p1*n1) + (p2*n2))/(n1 + n2) 
    z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/n1 + 1/n2)) # z-statistic
    if (z >0) {
            p <- 2 * pnorm(z, lower.tail = FALSE)
          }
          if (z < 0) {
            p <- 2 * pnorm(z, lower.tail = TRUE)
          }
    
    if(alternative == "greater"){
      p <- pnorm(z, lower.tail = FALSE)
    }
    if(alternative == "lower"){
      p <- pnorm(z, lower.tail = TRUE) 
    }
    
  #values for confidence interval
  upper <- (p1 - p2) + (qnorm(1 - 0.5*1 - (conf.level))) * (sqrt((p1*(1-p1)/n1) + (p2 * (1-p2)/n2)))
  lower <- (p1 - p2) - (qnorm(1 - 0.5*1 - (conf.level))) * (sqrt((p1*(1-p1)/n1) + (p2 * (1-p2)/n2)))
  ci <- c(lower, upper)
  
  #returning the statistics: 
  return("two-sample z prop-test")
    CI <- conf.level
    CI
    Alternative <- alternative
    Alternative
    z-value <- as.numeric(z)
    p-value <- as.numeric(p)
  }
}
```

This runs. I'll take that as a win for now. But, I'm scared to test it. 

## Question 2:

The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

1. Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

2. Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

3. Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.

4. Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?

5. Looking at your two models, which do you think is better? Why?

### Calling the dataset: 

```{r question2}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
f <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(f)
```

### Regression Models:

```{r regression-model}
f <- na.omit(f)
longevity <- f$MaxLongevity_m
brain_size <- f$Brain_Size_Species_Mean

beta1 <- cor(longevity,brain_size)*sd(longevity)/sd(brain_size)
beta1

beta0 <- mean(longevity) - beta1 * mean(brain_size)
beta0

#Therefore, the equation is 307.7522 = y - 0.8789037x

g <- ggplot(data = f, aes(x = brain_size, y = longevity))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g #not able to figure out the geom_text thing yet
```
```{r log-regression}
log_longevity <- log(f$MaxLongevity_m)
log_brain_size <- log(f$Brain_Size_Species_Mean)

beta1 <- cor(log_longevity,log_brain_size)*sd(log_longevity)/sd(log_brain_size)
beta1

beta0 <- mean(log_longevity) - beta1 * mean(log_brain_size)
beta0

#Therefore, the equation is 5.086498 = y - 0.2028543x

g <- ggplot(data = f, aes(x = log_brain_size, y = log_longevity))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g #not able to figure out the geom_text thing yet
```
### Confidence Intervals:

```{r calc-CI}
model <- lm(longevity ~ brain_size, data = f)
beta1_ci <- confint(model, level = 0.90)
beta1_ci
summary(model) #based on these results, the values are significant at 0.1 (and even higher)

#for the log-regression model
model_log <- lm(log_longevity ~ log_brain_size, data = f)
beta1_log_ci <- confint(model_log, level = 0.90)
beta1_log_ci
summary(model_log) #based on these results, the values are significant at 0.1 (and even higher)
```

I wasn't sure how to actually do a statistical test to see if beta1 is different from 0, but I used the summary function. I'm not sure if that is correct? 

I'm not sure what is meant by 'interpret the point estimate of the slope of beta1' 

### Prediction Values: 

```{r predictions_non_long}
#For the non-log model, the equation 307.7522 = y - 0.8789037x
lower_longevity <- 260.1112167+0.5928154*(brain_size) #values from calling beta1
lower_longevity #values of longevity based on the lower end of the CI

upper_longevity <- 355.393214 + 1.164992*(brain_size)
upper_longevity #values of longevity based on the upper end of the CI

#Prediction intervals:
pred_int <- predict(model, interval = "prediction", level = 0.90)
pred_int

intervals <- as.data.frame(pred_int)
intervals <- cbind(intervals, lower_longevity)
intervals <- cbind(intervals, upper_longevity)
intervals <- cbind(intervals, brain_size)
intervals <- cbind(intervals, longevity)

#Plotting: 
g <- ggplot(intervals, aes(x = brain_size, y= longevity))+ 
  geom_point()+
  geom_line(aes(x = brain_size, y = fit, color = "black"))+
  geom_line(aes(brain_size, lwr, color = "green"))+
  geom_line(aes(brain_size, upr, color = "green"))+
  geom_line(aes(brain_size, lower_longevity, color = "red"))+
  geom_line(aes(brain_size, upper_longevity, color = "red"))

#I need to add a legend but black is the best fit, green is the prediction intervals and red are the CIs

g  
```
```{r log_reg_predictions}

beta1_log_ci

#For the non-log model, the equation 5.086498 = y - 0.2028543x
log_lower_longevity <- 4.7030309+0.1172695*(log_brain_size) #values from calling beta1_log_ci
log_lower_longevity #values of log_longevity based on the lower end of the CI

log_upper_longevity <- 5.4699646 + 0.2884391*(log_brain_size)
log_upper_longevity #values of log_longevity based on the upper end of the CI

#Prediction intervals:
pred_int_log <- predict(model_log, interval = "prediction", level = 0.90)
pred_int_log

intervals_log <- as.data.frame(pred_int_log)
intervals_log <- cbind(intervals_log, log_lower_longevity)
intervals_log <- cbind(intervals_log, log_upper_longevity)
intervals_log <- cbind(intervals_log, log_brain_size)
intervals_log <- cbind(intervals_log, log_longevity)

#Plotting: 
g <- ggplot(intervals_log, aes(x = log_brain_size, y= log_longevity))+ 
  geom_point()+
  geom_line(aes(x = log_brain_size, y = fit, color = "black"))+
  geom_line(aes(log_brain_size, lwr, color = "green"))+
  geom_line(aes(log_brain_size, upr, color = "green"))+
  geom_line(aes(log_brain_size, log_lower_longevity, color = "red"))+
  geom_line(aes(log_brain_size, log_upper_longevity, color = "red"))

#I need to add a legend but black is the best fit, green is the prediction intervals and red are the CIs

g  
```
### Point Estimate for brain size = 800 gms 

```{r point_estimates}
point_estimate <- 307.7522 + 0.8789037*800
point_estimate #1010.875 months

log_point_estimate <- 5.086498 + 0.2028543*800 #I'm not sure what to do with this. 

```
Overall, I don't these are bad models to predict the longevity accurately. Brain_size is considered a constant which is used to predict longevity, which I don't think is a bad assumption to make. 
I do think that the log transformed model is better as it evens out the data a bit (takes care of the extreme values).

## Challenges I faced:

1. Making the function made me cry. I kept questioning if I knew what I was doing. Probably not. 

2. The function is so long and inefficient. I didn't have time to fix that. And I'm not sure if I've made the loops correctly to begin with. 

3. The geom_text function is so confusing. I still can't figure it out. 
